---
title: "Day1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Notes from Day 1 of the Programming and Analyzing R Workshop

setwd("~/Dropbox/ProgrammingAnalyzinginR/ProgrammingAnalyzinginR") #not necessary because the project will always open in the right directory
library(foreign) #call package foreign in order to use read.spss
Exp1 <- read.spss("DATA_Exp1_compact_correct67_outliers_1.sav", to.data.frame = TRUE) #read in data

save(Exp1, file="Exp1.rda") #to save data as an R data object/file. Rda file is R data file, it can remember al variable names, rds is also R data file but it can only contain only 1 variable. Don't know why this is necessary but this is something new I've learned today. 

load("Exp1.rda") #to load data

Exp1$PPNo <- c(1:154) #attach/create a new variable to the data object Exp1 

#Subsetting

subset_data <- dataframe[dataframe$Year > 1945, "France", drop = FALSE] #Without drop = FALSE then the output will be just a vector. With drop = FALSE the output will be still a dataframe. This is only a note from the lecture of Wouter. I do not have the data to run this.

#Using subsetting in statistical tests.

postwar <- data[data$Year > 1945,]
t.test(postwar$France, postwar$U.S., paired = TRUE) # t-test for the difference between France & US after 1945
with(data[data$Year > 1945, t.test(France, U.S.,paired = TRUE)])  # another way to do this t-test

#You can also use subsetting to do winsorizing

#First calculate 3sd & mean

#Then calculate the winsorized point = mean + 3sd

#Then assign the values that are larger than 3sd with the winsorized point using "="







```

## R Markdown

This is an R Markdown  ßdocument. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
